systemfit: A Package for Estimating\\ Systems of Simultaneous Equations in \proglang{R}

### 2007-06-06 paper/trunk/JSS/response_A.tex

\subsection{Abstract/Introduction}
%   o abstract/introduction
% 
%     Both, abstract and introduction, talk rather generally about
%     systems of equations reviewing rather well-known theoretical
%     results. I would suggest to revise both, focusing on what
%     *can* be done *in R* rather than giving a tutorial what *should*
%     be done *in general*. Of course, it is ok to review some theory,
%     however, practitioners should know what they are doing and the
%     paper should mainly teach them how to do it in R.
%
Both, abstract and introduction, now focus much more
on the capabilities of the \pkg{systemfit} package
rather than on the general theory of estimating systems of equations.

\subsection{Estimation Methods (2.1)}
%   o estimation methods (2.1)
% 
%     The description of the estimation methods can be shortened considerably
%     and should focus on the unifying properties of the different estimators.
%     All methods from OLS through to standard 3SLS comply with formulas
%     (12) and (13), only with differing choices for Omega (s*I, diagonal or
%     full Omega) and X (original X vs \hat X from instruments). The exposition
%     could be streamlined considerably if this structure were exploited.
% 
%     Terminology for the different estimators also varies between various
%     communities using systems of equations models. The paper should address
%     this. For example, in econometrics the GLS estimator would be an estimator
%     with known Omega while FGLS (feasible GLS) is the estimator based on
%     an estimated Omega (see Greene, 2003). The current discussion makes this
%     even more problematic because Omega is treated as known throughout
%     Section 2. The discussion of the estimators needs to be combined with
%     discussion of the estimators for Omega.
%
The paper no longer treats $\Omega$ as known in sections~2.1 and~2.2.
The fact that $\Sigma$ and $\Omega$ are unknown and have to be estimated
is no longer °hidden° in section 2.3,
but is mentioned in sections~2.1 and~2.2 now.

%     For the varying 3SLS estimators, it would help to point out why/when
%     these are needed at the beginning rather than at the end of the 3SLS
%     section. 
%
We think that it is better
to present the review of the varying 3SLS estimators
at the end rather than at the beginning of the 3SLS section,
because it is easier for the reader to understand this review
if the compared estimation methods have been described before.

\subsection{Imposing Linear Restrictions (2.2)}
%   o imposing linear restrictions (2.2)
% 
%     I found this section somewhat confusing. At first, it should be made
%     clear that this is about estimating the coefficients under linear
%     restrictions (and not inference or something else). Then, it needs
%     to be explained more concisely that all linear restrictions are of
%     type (33) and that the rest of the discussion is about different
%     computational algorithms for obtaining the estimates under restriction.
%     Again, the exposition on pages 7-9 can be improved (and shortened)
%     if the unifying structure is exploited. Also the notation can be
%     streamlined concerning beta, \hat beta, beta^0, beta^* etc.
%
\subsection{systemfitClassic (4.3)}
%   o systemfitClassic (4.3)
% 
%     First, this function deserves a better name: it is long, unintuitive
%     and does not reflect the underlying conceptual tools. If the requirement
%     of using the same formula for all equations would be relaxed (but
%     still supported) a conceivable name would be sur() for SUR type 
%     models. But it would also be conceivable to stick to the current
%     requirement of a single formula (then, still a better name would
%     be needed) and to set up a second wrapper sur() etc. In any case,
%     the function (its name, arguments and output) should reflect what
%     is conceptually intuitive for this kinds of models (for the communities
%     using it).
%     Second, the interface should also be re-designed: the standard in R is
%     that the formula/s is/are the first argument (which would also be more
%     convenient for systemfit()), e.g.,
%       function(formula, method = "FGLS", data = ..., ...)
%     Currently, the data argument seems to be mandatory (opposed to the
%     paper documentation). The specification of eqnVar and timeVar via
%     strings seems to be uncommon and suboptimal. Probably some kind of
%     formula would be preferable. Maybe a formula similar to those used in
%     nlme could be employed? If more than one formula is specified, one has
%     to pay attention that no problems are used by different patterns of
%     missingness in differing equations.
%
\subsection{Testing Linear Restrictions (4.4)}
%   o testing linear restrictions (4.4)
% 
%     The package introduces several functions that are somewhat orthogonal
%     to exisiting methodology. A single method (with "test" argument) for
%     the linear.hypothesis() function from "car" might seem more appropriate
%     here. The Wald test can already now be reproduced using the default
%     linear.hypothesis() method (except for formula formatting)
%       linear.hypothesis(fitsur, Rmat, qvec)    
%     If the coefficient naming would be modified (see below), the convenience
%     interface for specifying the restrictions in linear.hypothesis() could
%     also be re-used.    
%     Also two of the functions look like methods to the generics lrtest() and
%     waldtest() in package "lmtest", although this is not the case. Furthermore,
%     some more general method of comparing various fitted models would be nice,
%     e.g., an anova() method or methods to waldtest()/lrtest() although the 
%     latter two seem to be under development.
%     The output formatting of the current function is also not very elegant,
%     a new class along with its print method is introduced for each test. It
%     would be much easier to re-use "htest" or (probably better) "anova" as
%     classes for this.
%
\subsection{Organize Internals of systemfit}
%   o organize internals of systemfit
% 
%     The internals of a fitted "systemfit" object could be re-organized. Currently,
%     there are a lot of single letter elements - more standard naming as in "lm"
%     objects etc. would be more desirable.
We have renamed single letter elements (not finished yet).

### 2007-06-06 paper/trunk/systemfit_estimation.tex
The paper no longer treats Omega as known in sections 2.1 and 2.2; the fact that Sigma and Omega are unknown and have to be estimated is no longer "hidden" in section 2.3, but is mentioned in sections 2.1 and 2.2 now; added a note in the response to reviewer A that this has been changed in the paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@714 591db2bd-e0a0-4d22-8c19-2192dbb7da31

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y .
   \label{eq:ols}
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1} ,
   \label{eq:olsCovSameSigma}
\end{equation}
where $\sigma^2 = \E \left[ u_{it}^2 \right] \, \forall \, i$.
This assumes that the disturbances of all equations have the
same variance.

If the disturbance terms of the individual equations
are allowed to have different variances,
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{OLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} ,
   \label{eq:olsCovSingleSigma}
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.

@@

The Weighted Least Squares (WLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{WLS} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
Like the OLS estimates these estimates are only efficient
if the disturbance terms are not contemporaneously correlated.
The covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{WLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
\end{equation}
If no cross-equation parameter restrictions are imposed,
the parameter estimates are equal to the OLS estimates.

@@

When the disturbances are contemporaneously correlated, a Generalized 
Least Squares (GLS) estimation leads to efficient parameter estimates.
In this case, the GLS estimator is generally called °Seemingly Unrelated Regression°
(SUR) estimator \citep{zellner62}.
It should be noted that while an unbiased OLS or WLS estimation requires only that 
the regressors and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i ' X_i \right] = 0 \; \forall \; i )$,
a consistent SUR estimation requires that all disturbance terms and all 
regressors are uncorrelated
$( \E \left[ u_i ' X_j \right] = 0 \; \forall \; i, j )$.

The SUR estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{SUR} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
   \label{eq:sur}
\end{equation}
where $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
And the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
   \label{eq:surCov}
\end{equation}

@@

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X .
\end{equation}
At the second stage the unbiased two-stage least squares estimates
of $\beta$ are obtained by
\begin{equation}
   \widehat{\beta}_{2SLS} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
   \widehat{X}' y .
   \label{eq:beta2sls}
\end{equation}
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{2SLS} \right] = \sigma^2 \left( \widehat{X}'
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSameSigma}
\end{equation}
where $\sigma^2 = \E \left[ u_{it}^2 \right]$.
If the disturbance terms of the individual equations
are allowed to have different variances, 
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSingleSigma}
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.

@@

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
The covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} .
\end{equation}

@@

If the regressors are correlated with the disturbances 
($\E \left[ u_i' X_i \right] \neq 0$) and
the disturbances are contemporaneously correlated, 
a Generalized Least Squares (GLS) version of the two-stage least squares
estimation leads to consistent and efficient estimates.
This estimation procedure is generally called °Three-stage Least
Squares° (3SLS) \citep{zellner62b}.

The standard 3SLS estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
   \label{eq:3slsGls}
\end{equation}
where $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
Its covariance matrix is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} .
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] .
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
One of these approaches based on instrumental variable estimation
(3SLS-IV) is
\begin{equation}
   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \Omega^{-1} X
   \right)^{-1} \widehat{X}' \Omega^{-1} y .
   \label{eq:3slsIv}
\end{equation}
The covariance matrix of this 3SLS-IV estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
   X \right)^{-1} .
\end{equation}
Another approach based on the Generalized Method of Moments (GMM)
estimator (3SLS-GMM) is
\begin{equation}
   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \Omega H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y .
   \label{eq:3slsGmm}
\end{equation}
The covariance matrix of the 3SLS-GMM estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-GMM} \right] =
   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1} .
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is
\begin{equation}
   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \Omega^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \Omega^{-1} 
   H \left( H' H \right)^{-1} H' y .
   \label{eq:3slsSchmidt}
\end{equation}
The covariance matrix of this estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} .
\end{equation}
The econometrics software \proglang{EViews} uses
\begin{equation}
   \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
   \label{eq:3slsEViews}
\end{equation}
where $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
as defined by (\ref{eq:beta2sls}).
\proglang{EViews} uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate the covariance matrix of the 3SLS estimator.

>>>>

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y .
   \label{eq:ols}
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{OLS} \right] = \sHat^2 \left( X'X \right)^{-1} ,
   \label{eq:olsCovSameSigma}
\end{equation}
where $\sHat^2$ is the estimated variance of the residuals.
This assumes that the disturbances of all equations have the
same variance $(\E [ u_{it}^2 ] = \sigma^2 \, \forall \, i)$.

If the disturbance terms of the individual equations
are allowed to have different variances,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{OLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} ,
   \label{eq:olsCovSingleSigma}
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbence term in the $i$th equation.

@@

The Weighted Least Squares (WLS) estimator allows for different variances
of the disturbance terms in the different equations
$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j$,
where $\sigma_i^2 = \E [ u_{it}^2 ])$.
The WLS estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{WLS} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.%
\footnote{%
Theoretically, one should use the variances of the true residuals.
However, they are not known in most empirical applications.
Therefore, true variances are generally replaced by estimated variances
that are calculated from the residuals of a first-step OLS estimation.
}
Like the OLS estimates, these estimates are only efficient
if the disturbance terms are not contemporaneously correlated.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{WLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
\end{equation}
If no cross-equation parameter restrictions are imposed,
the parameter estimates are equal to the OLS estimates.

@@

When the disturbances are contemporaneously correlated, a Generalized 
Least Squares (GLS) estimation leads to efficient parameter estimates.
In this case, the GLS estimator is generally called °Seemingly Unrelated Regression°
(SUR) estimator \citep{zellner62}.
Since the true covariance matrix of the disturbance terms
is generally unknown,
a Feasible Generalized Least Squares (FGLS) estimation,
which is based on an estimated covariance matrix of the disturbance terms,
is the textbook solution for this problem.
It should be noted that while an unbiased OLS or WLS estimation requires only that
the regressors and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i ' X_i \right] = 0 \; \forall \; i )$,
a consistent SUR estimation requires that all disturbance terms and all 
regressors are uncorrelated
$( \E \left[ u_i ' X_j \right] = 0 \; \forall \; i, j )$.

The SUR estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{SUR} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
   \label{eq:sur}
\end{equation}
where $\OHat = \SHat \otimes I$ and
$\SHat$ is the estimated covariance matrix of the disturbance terms.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{SUR} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
   \label{eq:surCov}
\end{equation}

@@

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X .
\end{equation}
At the second stage, the unbiased two-stage least squares estimates
of $\beta$ are obtained by
\begin{equation}
   \widehat{\beta}_{2SLS} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
   \widehat{X}' y .
   \label{eq:beta2sls}
\end{equation}
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{2SLS} \right] = \sHat^2 \left( \widehat{X}'
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSameSigma}
\end{equation}
where $\sHat^2$ is the estimated variance of the disturbance terms.
If the disturbance terms of the individual equations
are allowed to have different variances, 
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \OHat^{-1}
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSingleSigma}
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.

@@

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \OHat^{-1} y ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \OHat^{-1}
   \widehat{X} \right)^{-1} .
\end{equation}

@@

If the regressors are correlated with the disturbances 
($\E \left[ u_i' X_i \right] \neq 0$) and
the disturbances are contemporaneously correlated, 
a Feasible Generalized Least Squares (FGLS) version of the two-stage least squares
estimation leads to consistent and asymptotically efficient estimates.
This estimation procedure is generally called °Three-stage Least
Squares° (3SLS) \citep{zellner62b}.

The standard 3SLS estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \OHat^{-1} y ,
   \label{eq:3slsGls}
\end{equation}
where $\OHat = \SHat \otimes I$ and
$\SHat$ is the estimated covariance matrix of the disturbance terms.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \OHat^{-1}
   \widehat{X} \right)^{-1} .
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] .
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
One of these approaches based on instrumental variable estimation
(3SLS-IV) is
\begin{equation}
   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \OHat^{-1} X
   \right)^{-1} \widehat{X}' \OHat^{-1} y .
   \label{eq:3slsIv}
\end{equation}
An estimator of the covariance matrix of the estimated 3SLS-IV parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \OHat^{-1}
   X \right)^{-1} .
\end{equation}
Another approach based on the Generalized Method of Moments (GMM)
estimator (3SLS-GMM) is
\begin{equation}
   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \OHat H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \OHat H \right)^{-1} H' y .
   \label{eq:3slsGmm}
\end{equation}
An estimator of the covariance matrix of the estimated 3SLS-GMM parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{3SLS-GMM} \right] =
   \left( X' H \left( H' \OHat H \right)^{-1} H' X \right)^{-1} .
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is
\begin{equation}
   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \OHat^{-1} 
   H \left( H' H \right)^{-1} H' y .
   \label{eq:3slsSchmidt}
\end{equation}
An estimator of the covariance matrix of these estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
   \left( \widehat{X}' \OHat^{-1}  \widehat{X} \right)^{-1}
   \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
   H \left( H' H \right)^{-1} H' \OHat^{-1} \widehat{X}
   \left( \widehat{X}' \OHat^{-1}  \widehat{X} \right)^{-1} .
\end{equation}
The econometrics software \proglang{EViews} uses
\begin{equation}
   \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \OHat^{-1} \widehat{X} \right)^{-1}
   \widehat{X}' \OHat^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
   \label{eq:3slsEViews}
\end{equation}
where $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
as defined by (\ref{eq:beta2sls}).
\proglang{EViews} uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate an estimator of the covariance matrix of the estimated parameters.
